---
title: "ggplottimeseries: Visualisation of decomposed time series using ggplot"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This package contains functions that allow users to plot decomposed time series data in ggplot. 

## Installation

To install this package, you can use the following codes: 

```{r message=FALSE}

devtools:: install_github("brisneve/ggplottimeseries")
library(ggplottimeseries)

```

## Functions

**Decompose time series data into its components (i.e. trends, seasonal, random) and collate them into a data frame.**

To demonstrate this, we will use the function *dts1*  to decompose the daily sea surface temperature data into its three components. This dataset has 2142 observations on the following 2 variables: 

1. date - a vector of time of observation (class: data, format: YYYY-MM-DD)
2. sst - sea surfae temperature data (class: numeric)

The *dts1* function requires the following inputs: 

1. x - dates of the observations 
2. y - observations 
3. z - frequency of the obserbation over a period of time (e.g. year) 
4. type - type of timeseries (additive or multiplicative) 

Try the codes below and explore the output data frame which contains the components of the time series:

```{r message=FALSE}
data(sst)

x <- sst$date
y <- sst$sst
z <- 365.25 #number of days in a year

df <- dts1(x,y,z, type = "additive")

head(df)


```


Alternatively, if you have a time series-class data you can use the function *dts2*. See how this function works using the co2 dataset that contains a time series of atmospheric concentration of CO2. This function creates a dataframe of the time series components similar to the *dts2*. 

```{r message=FALSE}

data(co2)

df <- dts2(co2, type ="additive")

head(df)

```

We now have a data frame that we could plot using ggplot. This package offers suite of functions to plot the decomposed time series data in ggplot and can be integrated with ggplot functions.

**Visualise the observed, trend, seasonal, and random components of time series into one figure**  


```{r message=FALSE}


#plots decomposed time series into one figure
ggdecompose(df)+
  xlab("Date")+
  ylab("Atmospheric Concentration of CO2")

```








**Visualise the observed, trend, seasonal, and random components of time series into one figure**   separate figures (ggobserve, ggtrend, ggseason, ggrandom, respectively). These functions also plots detrended and deseasonalised time series (ggdetrend and ggseason, respectively). These can be integrated with ggplot functions.

